Transactional Memory Concurrency Verification with Landslide

### 2018-03-30 htm/for-garth/paper.tex
update forg-arth

\[
\begin{tabular}{ll|ll|ll}
	\multicolumn{2}{c|}{\hilight{lavender}{Thread 1}} &
	\multicolumn{2}{c|}{\hilight{seafoam}{Thread 2}} &
	\multicolumn{2}{c}{\hilight{salmon}{Thread 3}} \\
	\hline
	\hilight{lavender}{$1$} & \hilight{lavender}{\texttt{\_xbegin();}} & & \\
	%
	& & \hilight{seafoam}{$1$} & \hilight{seafoam}{\texttt{\_xbegin();}} \\
	& & \hilight{seafoam}{$2$} & \hilight{seafoam}{\texttt{x++;}} \\
	& & \hilight{seafoam}{$3$} & \hilight{seafoam}{\texttt{\_xend();}} \\
	%
	\hilight{lavender}{$5$} & \hilight{lavender}{\texttt{mutex\_lock(m);}} & & \\
	\hilight{lavender}{$6a$} & \hilight{lavender}{\texttt{temp <- x;}} & & \\
	%
	& & & & \hilight{salmon}{$1$} & \hilight{salmon}{\texttt{\_xbegin();}} \\
	& & & & \hilight{salmon}{$2$} & \hilight{salmon}{\texttt{x++;}} \\
	& & & & \hilight{salmon}{$3$} & \hilight{salmon}{\texttt{\_xend();}} \\
	%
	\hilight{lavender}{$6b$} & \hilight{lavender}{\texttt{temp <- temp + 1;}} & & \\
	\hilight{lavender}{$6c$} & \hilight{lavender}{\texttt{x <- temp;}} & & \\
	\hilight{lavender}{$7$} & \hilight{lavender}{\texttt{mutex\_unlock(m);}} & & \\
\end{tabular}
\]

Prior work \cite{htm-mario} proposed the idiom shown in Figure~\ref{fig:htm-fixed}
to exclude interleavings which violate the failure path's atomicity.
Thread 1 would set the {\tt prevent\_transactions} flag across its {\tt x++} accesses,
and
Thread 3 would have to check it and trigger an abort. %(Figure \ref{fig:htm-fixed} lines 2-3).
%which shows that correctly synchronizing even the simplest transactions may be surprisingly difficult or complex.
%further motivating our research.

@@

\section{Implementation}

@@

\subsection{Bugs}

@@

@@

\begin{figure}[t]
	\begin{center}
		\begin{tabular}{l}
		%\texttt{void count() \{} \\
		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
		\texttt{\flow{while} (\_retry);} \\
		\texttt{\flow{if} (\call{\_xbegin}() == \const{SUCCESS}) \{} \\
		\texttt{~~~~\hilight{brickred}{tie}(\_root,inserted) = \call{\_insert}(\_root,n); } \\
		\texttt{~~~~\call{\_xend}();} \\
		\texttt{\} \flow{else} \{} \\
		\texttt{~~~~\call{pthread\_mutex\_lock}(\&\_tree\_lock);} \\
		\texttt{~~~~\_retry = \const{true};} \\
		\texttt{~~~~\hilight{brickred}{tie}(\_root,inserted) = \call{\_insert}(\_root,n); } \\
		\texttt{~~~~\_retry = \const{false};} \\
		\texttt{~~~~\call{pthread\_mutex\_unlock}(\&\_tree\_lock);} \\
		\texttt{\}} \\
		\end{tabular}
	\end{center}
	%\caption{mario man had a very large house. or a castle.}
	\caption{Unmodified code from {\tt htmavl.cpp} showing the previously-unknown bug found by Landslide.
		The transaction path fails to check {\tt \_retry},
		leading to data races and corruption just as in {\tt htm1}.
		}
	\label{fig:avlbug}
\end{figure}

\subsection{Verification}

Landslide proved several of our test configurations correct, as shown in Table~\ref{tab:verifs}
%With no bugs to find, comparing the different testing modes against each other is meaningless;
%we simply present their state space sizes and runtime (using {\tt -M}) in Table~\ref{tab:verifs}.
Several larger tests we were unable to complete. %within several hours.
% バレちゃった
%before the submission deadline are also shown.
%Their completion times are listed as {\em unk.}
These are indicated with $\dagger$ %\ETAdag{}
and their listed ETAs and state space sizes represent Landslide's estimate %at the moment of timeout.
after a timeout of 1 hour
(hopefully enough for data-race PPs to saturate and estimates to stablize,
although note the two estimate types use different algorithms so may disagree significantly).

>>>>

\[
\begin{tabular}{ll|ll|ll}
	\multicolumn{2}{c|}{\hilight{lavender}{Thread 1}} &
	\multicolumn{2}{c|}{\hilight{seafoam}{Thread 2}} &
	\multicolumn{2}{c}{\hilight{salmon}{Thread 3}} \\
	\hline
	\hilight{lavender}{$1$} & \hilight{lavender}{\texttt{\_xbegin();}} & & \\
	%
	& & \hilight{seafoam}{$1$} & \hilight{seafoam}{\texttt{\_xbegin();}} \\
	& & \hilight{seafoam}{$2$} & \hilight{seafoam}{\texttt{x++;}} \\
	& & \hilight{seafoam}{$3$} & \hilight{seafoam}{\texttt{\_xend();}} \\
	%
	& \hilight{lavender}{\em transaction fails} & & & \\
	\hilight{lavender}{$5$} & \hilight{lavender}{\texttt{mutex\_lock(m);}} & & \\
	\hilight{lavender}{$6a$} & \hilight{lavender}{\texttt{temp <- x;}} & & \\
	%
	& & & & \hilight{salmon}{$1$} & \hilight{salmon}{\texttt{\_xbegin();}} \\
	& & & & \hilight{salmon}{$2$} & \hilight{salmon}{\texttt{x++;}} \\
	& & & & \hilight{salmon}{$3$} & \hilight{salmon}{\texttt{\_xend();}} \\
	%
	\hilight{lavender}{$6b$} & \hilight{lavender}{\texttt{temp <- temp + 1;}} & & \\
	\hilight{lavender}{$6c$} & \hilight{lavender}{\texttt{x <- temp;}} & & \\
	\hilight{lavender}{$7$} & \hilight{lavender}{\texttt{mutex\_unlock(m);}} & & \\
\end{tabular}
\]

Prior work \cite{htm-mario} proposed the idiom shown in Figure~\ref{fig:htm-fixed}
to exclude interleavings which violate the failure path's atomicity such as shown above.
Thread 1 would set the {\tt prevent\_transactions} flag across its {\tt x++} accesses,
and
Thread 3 would have to check it and trigger an abort. %(Figure \ref{fig:htm-fixed} lines 2-3).
%which shows that correctly synchronizing even the simplest transactions may be surprisingly difficult or complex.
%further motivating our research.

@@

\newpage
\section{Implementation}

@@

\subsection{Bugs}
\label{sec:bugs}

@@

		A parameter of $K,N$ denotes $K$ threads each running the test logic for $N$ iterations.

@@

\begin{figure}[t]
	\begin{center}
		\begin{tabular}{l}
		%\texttt{void count() \{} \\
		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
		\texttt{\flow{while} (\_retry);} \\
		\texttt{\flow{if} (\call{\_xbegin}() == \const{SUCCESS}) \{} \\
		\texttt{~~~~\hilight{brickred}{tie}(\_root,inserted) = \call{\_insert}(\_root,n); } \\
		\texttt{~~~~\call{\_xend}();} \\
		\texttt{\} \flow{else} \{} \\
		\texttt{~~~~\call{pthread\_mutex\_lock}(\&\_tree\_lock);} \\
		\texttt{~~~~\_retry = \const{true};} \\
		\texttt{~~~~\hilight{brickred}{tie}(\_root,inserted) = \call{\_insert}(\_root,n); } \\
		\texttt{~~~~\_retry = \const{false};} \\
		\texttt{~~~~\call{pthread\_mutex\_unlock}(\&\_tree\_lock);} \\
		\texttt{\}} \\
		\end{tabular}
	\end{center}
	%\caption{mario man had a very large house. or a castle.}
	\caption{Unmodified code from {\tt htmavl.hpp} showing the previously-unknown bug found by Landslide.
		The transaction path fails to check {\tt \_retry},
		leading to data races and corruption just as in {\tt htm1}.
		}
	\label{fig:avlbug}
\end{figure}

\newpage
\subsection{Verification}

Landslide proved several of our test configurations correct, as shown in Table~\ref{tab:verifs}
%With no bugs to find, comparing the different testing modes against each other is meaningless;
%we simply present their state space sizes and runtime (using {\tt -M}) in Table~\ref{tab:verifs}.
Several larger tests we were unable to complete. %within several hours.
% バレちゃった
%before the submission deadline are also shown.
%Their completion times are listed as {\em unk.}
These are indicated with $\dagger$ %\ETAdag{}
and their listed ETAs and state space sizes represent Landslide's estimate %at the moment of timeout.
after a timeout of 1 hour
(hopefully enough for data-race PPs to saturate and estimates to stablize).
Note that the two estimate types use different underlying algorithms, so may disagree significantly,
such as in {\tt avl\_fixed}$(2,2)$.

While we verified {\tt avl\_fixed} for 2 threads and 1 iteration,
the buggy {\tt avl\_insert} actually required higher parameters ($2,2$ or $3,1$) to expose the bug in the first place,
either of which produced too large of a state space under {\tt avl\_fixed},
so we were unable to formally verify that our fix actually removed the bug.
Likewise, {\tt map\_basic}, which includes a concurrent map-resizing step,
could not be verified even under the minimum parameters ($2,1$);
at best we were able to verify a modified version {\tt map\_basicer}
which started with a pre-sized map to test inserts only.
Possible approaches to bring the sizes of these state spaces within reach
include testing the synchronization mechanism and the data structure logic separately, as discussed in Section~\ref{sec:bugs},
upgrading Landslide's exploration algorithms to achieve better reduction,
\cite{optimal-dpor,mcr}.
and identifying new types of equivalent interleavings under the new concurrency model \cite{thesprop}.
